### 2023-12-15 manuscript/manuscript.tex
added multiple comparison by year and town graphs

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
% for space filling
\usepackage{lipsum}
% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}
\usepackage{tabularx}
\usepackage{adjustbox}
\usepackage[utf8]{inputenc}

% double spacing
\usepackage{setspace}
\doublespacing

%% meta data

\title{Bridging Gaps: Investigating COVID-19's Influence on Health Disparities in Connecticut}
\author{Delia Lin\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}
\maketitle

\begin{abstract}
  Social determinants of health (SDoH) are the conditions in which people are born, 
  grow, live, work, and age, significantly influencing their overall health and well-being. 
  These determinants include socioeconomic status, education, access to healthcare, and the 
  physical environment. Understanding the interactions of these elements will be essential for 
  addressing health disparities and developing more effective public health policies and interventions.
\end{abstract}

\section{Introduction}\label{sec:intro}

Current research focuses on how social determinants of health (SDoH) plays a  massive
impact on one's health; it is estimated that 80 percent of a population's health outcomes are 
dictated by SDoH \citep{HOOD2016129}. Often, SDoH, when referring to an individual, can result in racial 
disparities in care when looking at a population\citep{Monroe2023-uq}. It has been shown that major inefficiencies
in the health system are attributed to overlooked prevention opportunities and unequal access
to care.\citep{Allin2014-xn}

Understanding the intricate interplay of these social determinants is crucial in addressing health disparities 
and developing effective public health policies and interventions. The COVID-19 pandemic has shed new light on 
these disparities, amplifying existing inequalities within various communities. This research topic gains paramount 
importance in the current context as it seeks to delve into the specific impact of COVID-19 on key social determinants 
of health in different counties and racial groups in Connecticut.

The existing literature underscores the pressing need for research in this area. Studies have shown that predominantly 
black counties in the United States experience significantly higher COVID-19 infection and mortality rates, emphasizing 
the racial disparities prevalent in healthcare outcomes. The pandemic has magnified these discrepancies, leading to 
mortality rates among historically marginalized minority communities that are 1.9 to 2.4 times higher compared to the 
general population \citep{Badalov2022-wt}. Additionally, inefficiencies in the healthcare system have been attributed 
to overlooked prevention opportunities and unequal access to care, necessitating a comprehensive examination 
of these social determinants in the context of the pandemic.

Despite the growing body of research on SDoH, there is a notable gap concerning the specific impact of COVID-19 on 
these determinants within diverse communities. This research aims to bridge this gap by comprehensively assessing how 
the pandemic has influenced key social determinants of health across various counties and racial groups in Connecticut. 
By identifying the specific ways in which different communities were affected, this study contributes valuable insights 
for targeted interventions, policy-making, and the development of equitable healthcare strategies.

% roadmap
The rest of the paper is organized as follows.

- The data will be presented in Section \ref{sec:data}

- The methods are described in Section \ref{sec:meth}

- The results are reported in Section \ref{sec:resu}

- A discussion concludes in Section \ref{sec:disc}

\section{Data}\label{sec:data}

Data was collected from The Agency for Healthcare Research and Quality (AHRQ).The dataset comprises 7 variables 
spanning a period of 4 years (2017, 2018, 2019, 2020) with observations across the 8 counties(Fairfield County, Hartford County, 
Litchfield County, Middlesex County, New Haven County, New London County, Tolland County, Windham County) in Connecticut. These variables 
encompass a total of 56 observations. The variables questions include housing, education level, income, insurance,
rehospitalization rates, food stamps usage, and population racial characteristics. The dataset includes a range of 
calculated percentages, median values, and raw observations, providing a holistic view of various factors affecting 
the communities in these counties.

\section{Methods}\label{sec:meth}

In this study, descriptive statistics is utilized to outline the total population, racial composition, 
education levels, and average rehospitalization rate across the 8 counties in Connecticut. Chi-squared analyses were 
conducted to assess the significance of the difference of the variables of median income, poverty level, health insurance, 
utilities access, and electronics access between counties and across the four years. Additional 
Tukey's HSD tests were conducted to determine the specific counties and years that have had significant differences within 
each of the variables for each county.

\section{Results}\label{sec:resu}

\ref{tab:Demographics1}
\ref{tab:Demographics2}

\ref{fig:Education by County Graph}
discuss, use lit to support why

\ref{fig:Social Needs by County Graph}
differences by year

\ref{tab:Multiple Comparison for Income}
the specific values for income

\ref{tab: Multiple Comparisons}
what discussion talks about

\begin{table}[h]
\label{tab:Demographics1}
  \caption{Demographics and Education Levels for Counties 1-4}
  \resizebox{\columnwidth}{!}{%
  \begin{tabular}{lccccc}
    \toprule
    \textbf{Category} & \textbf{Fairfield County} & \textbf{Hartford County} & \textbf{Litchfield County} & \textbf{Middlesex County} \\
    \midrule
    Total Population & 944,977 & 894,465.25 & 182,657.5 & 163,318.25 \\
    Race (Percent) & & & & \\
    \quad American Indian and Alaska Native & 0.24 & 0.3075 & 0.2025 & 0.195 \\
    \quad Asian & 5.2925 & 5.2925 & 1.9 & 3.0625 \\
    \quad Black or African American & 11.405 & 13.7025 & 1.83 & 5.385 \\
    \quad Native Hawaiian and Pacific Islander & 0.055 & 0.035 & 0 & 0.005 \\
    \quad White & 72.6325 & 70.67 & 92.6025 & 88.0875 \\
    Ethnicity (Percent) & & & & \\
    \quad Hispanic & 19.53 & 17.8275 & 6.15 & 6.12 \\
    \bottomrule
  \end{tabular}%
  }
\end{table}

\begin{table}[h]
  \label{tab:Demographics2}
  \caption{Demographics and Education Levels for Counties 5-8}
  \resizebox{\columnwidth}{!}{%
  \begin{tabular}{lcccc}
    \toprule
    \textbf{Category} & \textbf{New Haven County} & \textbf{New London County} & \textbf{Tolland County} & \textbf{Windham County} \\
    \midrule
    Total Population & 858,678 & 268,477.75 & 151,218.75 & 116,608.75 \\
    Race (Percent) & & & & \\
    \quad American Indian and Alaska Native & 0.1725 & 0.605 & 0.05 & 0.565 \\
    \quad Asian & 4.005 & 4.12 & 4.675 & 1.3675 \\
    \quad Black or African American & 13.34 & 5.8175 & 3.1075 & 2.33 \\
    \quad Native Hawaiian and Pacific Islander & 0.0225 & 0.025 & 0 & 0.015 \\
    \quad White & 73.2875 & 80.6175 & 88.025 & 88.8725 \\
    Ethnicity (Percent) & & & & \\
    \quad Hispanic & 17.885 & 10.5 & 5.4475 & 11.6375 \\
    \bottomrule
  \end{tabular}%
  }
\end{table}

\begin{figure}[tbp]
  \label{fig:Education by County Graph}
    \centering
    \includegraphics[width=\textwidth]{Education Levels by County.pdf}
    \caption{Education Levels by County.}
  \end{figure}

\begin{figure}[tbp]
  \label{fig:Social Needs by County Graph}
    \centering
    \includegraphics[width=\textwidth]{Multiple Comparisons by County.pdf}
    \caption{Social Needs Multiple Comparisons by County.}
  \end{figure}

  \begin{table}[htbp]
    \label{tab:Multiple Comparison for Income}
    \caption{Median Income Multiple Comparison by County}
    \centering
    \resizebox{\columnwidth}{!}{%
    \begin{tabular}{lllrrrr}
    \toprule
    & & & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{2}{l}{95\% Confidence} \\
    County 1(I) & County 2(J) & Mean Difference(I-J) & \multicolumn{1}{l}{Std. Error} & \multicolumn{1}{l}{Sig.} & \multicolumn{1}{l}{Lower Bound} & \multicolumn{1}{l}{Upper Bound} \\
    \midrule
    Fairfield County & Hartford County & 20565.50* & 1883.47 & \textbf{\textless 0.001} & 14327.60 & 26803.40 \\
    & Litchfield County & 14919.50* & 1883.47 & \textbf{\textless 0.001} & 8681.60 & 21157.40 \\
    & Middlesex County & 9671.75* & 1883.47 & \textbf{<.001} & 3433.85 & 15909.65 \\
    & New Haven County & 25662.75* & 1883.47 & \textbf{<.001} & 19424.85 & 31900.65 \\
    & New London County & 21456.50* & 1883.47 & \textbf{<.001} & 15218.60 & 27694.40 \\
    & Tolland County & 8705.00* & 1883.47 & \textbf{0.002} & 2467.10 & 14942.90 \\
    & Windham County & 28671.00* & 1883.47 & \textbf{<.001} & 22433.10 & 34908.90 \\
    \midrule
    Hartford County & Fairfield County & -20565.50* & 1883.47 & \textbf{<.001} & -26803.40 & -14327.60 \\
    & Litchfield County & -5646 & 1883.47 & 0.096 & -11883.90 & 591.90 \\
    & Middlesex County & -10893.75* & 1883.47 & \textbf{<.001} & -17131.65 & -4655.85 \\
    & New Haven County & 5097.25 & 1883.47 & 0.169 & -1140.65 & 11335.15 \\
    & New London County & 891 & 1883.47 & 1 & -5346.90 & 7128.90 \\
    & Tolland County & -11860.50* & 1883.47 & \textbf{<.001} & -18098.40 & -5622.60 \\
    & Windham County & 8105.50* & 1883.47 & \textbf{0.005} & 1867.60 & 14343.40 \\
    \midrule
    Litchfield County & Fairfield County & -14919.50* & 1883.47 & \textbf{<.001} & -21157.40 & -8681.60 \\
    & Hartford County & 5646 & 1883.47 & 0.096 & -591.90 & 11883.90 \\
    & Middlesex County & -5247.75 & 1883.47 & 0.145 & -11485.65 & 990.15 \\
    & New Haven County & 10743.25* & 1883.47 & \textbf{<.001} & 4505.35 & 16981.15 \\
    & New London County & 6537 & 1883.47 & 0.035 & 299.10 & 12774.90 \\
    & Tolland County & -6214.5 & 1883.47 & 0.051 & -12452.40 & 23.40 \\
    & Windham County & 13751.50* & 1883.47 & \textbf{<.001} & 7513.60 & 19989.40 \\
    \midrule
    Middlesex County & Fairfield County & -9671.75* & 1883.47 & \textbf{<.001} & -15909.65 & -3433.85 \\
    & Hartford County & 10893.75* & 1883.47 & \textbf{<.001} & 4655.85 & 17131.65 \\
    & Litchfield County & 5247.75 & 1883.47 & 0.145 & -990.15 & 11485.65 \\
    & New Haven County & 15991.00* & 1883.47 & \textbf{<.001} & 9753.10 & 22228.90 \\
    & New London County & 11784.75* & 1883.47 & \textbf{<.001} & 5546.85 & 18022.65 \\
    & Tolland County & -966.75 & 1883.47 & 0.999 & -7204.65 & 5271.15 \\
    & Windham County & 18999.25* & 1883.47 & \textbf{<.001} & 12761.35 & 25237.15 \\
    \midrule
    New Haven County & Fairfield County & -25662.75* & 1883.47 & <.001 & -31900.65 & -19424.85 \\
    & Hartford County & -5097.25 & 1883.47 & 0.169 & -11335.15 & 1140.65 \\
    & Litchfield County & -10743.25* & 1883.47 & <.001 & -16981.15 & -4505.35 \\
    & Middlesex County & -15991.00* & 1883.47 & <.001 & -22228.90 & -9753.10 \\
    & New London County & -4206.25 & 1883.47 & 0.368 & -10444.15 & 2031.65 \\
    & Tolland County & -16957.75* & 1883.47 & <.001 & -23195.65 & -10719.85 \\
    & Windham County & 3008.25 & 1883.47 & 0.747 & -3229.65 & 9246 \\
    \midrule
    New London County & Fairfield County & -21456.50* & 1883.47 & <.001 & -27694.40 & -15218.60 \\
    & Hartford County & -891 & 1883.47 & 1 & -7128.90 & 5346.90 \\
    & Litchfield County & -6537.00* & 1883.47 & 0.035 & -12774.90 & -299.10 \\
    & Middlesex County & -11784.75* & 1883.47 & <.001 & -18022.65 & -5546.85 \\
    & New Haven County & 4206.25 & 1883.47 & 0.368 & -2031.65 & 10444.15 \\
    & Tolland County & -12751.50* & 1883.47 & <.001 & -18989.40 & -6513.60 \\
    & Windham County & 7214.50* & 1883.47 & 0.016 & 976.60 & 13452.40 \\
    \midrule
    Tolland County & Fairfield County & -8705.00* & 1883.47 & 0.002 & -14942.90 & -2467.10 \\
    & Hartford County & 11860.50* & 1883.47 & <.001 & 5622.60 & 18098.40 \\
    & Litchfield County & 6214.50 & 1883.47 & 0.051 & -23.40 & 12452.40 \\
    & Middlesex County & 966.75 & 1883.47 & 0.999 & -5271.15 & 7204.65 \\
    & New Haven County & 16957.75* & 1883.47 & <.001 & 10719.85 & 23195.65 \\
    & New London County & 12751.50* & 1883.47 & <.001 & 6513.60 & 18989.40 \\
    & Windham County & 19966.00* & 1883.47 & <.001 & 13728.10 & 26203.90 \\
    \midrule
    Windham County & Fairfield County & -28671.00* & 1883.47 & <.001 & -34908.90 & -22433.10 \\
    & Hartford County & -8105.50* & 1883.47 & 0.005 & -14343.40 & -1867.60 \\
    & Litchfield County & -13751.50* & 1883.47 & <.001 & -19989.40 & -7513.60 \\
    & Middlesex County & -18999.25* & 1883.47 & <.001 & -25237.15 & -12761.35 \\
    & New Haven County & -3008.25 & 1883.47 & 0.747 & -9246.15 & 3229.65 \\
    & New London County & -7214.50* & 1883.47 & 0.016 & -13452.40 & -976.60 \\
    & Tolland County & -19966.00* & 1883.47 & <.001 & -26203.90 & -13728.10 \\
    \bottomrule
  \end{tabular}%
  }
\end{table}

\begin{table}[]
  \label{tab: Multiple Comparisons}
  \caption{Multiple Comparisons by Year and Town}
  \resizebox{\columnwidth}{!}{%
  \begin{tabular}{llrrrrl}
  \textbf{Social Need}   & \textbf{Comparison by} & \multicolumn{1}{l}{\textbf{DF}} & \multicolumn{1}{l}{\textbf{Sum of Squares}} & \multicolumn{1}{l}{\textbf{Mean Square}} & \multicolumn{1}{l}{\textbf{F Value}} & \textbf{P Value}           \\
  Median Income          & Year                   & 3                               & 158657606                                   & 52885869                                 & 95.57                                & <.0001                     \\
                         & Town                   & 7                               & 2618521106                                  & 374074444                                & 675.99                               & <.0001                     \\
  Food                   & Year                   & 3                               & 3.6450625                                   & 1.2150208                                & 10.08                                & 0.0003 \\
                         & Town                   & 7                               & 348.3382875                                 & 49.7626125                               & 412.72                               & <.0001                     \\
  Housing                & Year                   & 3                               & 1.4535125                                   & 0.4845042                                & 0.41                                 & 0.7506 \\
                         & Town                   & 7                               & 183.9498375                                 & 26.2785482                               & 21.99                                & <.0001                     \\
  Rehospitalization Rate & Year                   & 3                               & 0.00050938                                  & 0.00016979                               & 1.45                                 & 0.2578 \\
                         & Town                   & 7                               & 0.00427188                                  & 0.00061027                               & 5.2                                  & 0.0015 \\
  Uninsured Rate         & Year                   & 3                               & 2.6848375                                   & 0.89494583                               & 17.17                                & <.0001                     \\
                         & Town                   & 7                               & 92.8554875                                  & 13.26506964                              & 254.57                               & <.0001                    
  \end{tabular}%
  }
  \end{table}

\section{Discussion}\label{sec:disc}

The results of the analysis shed light on the intricate relationships between county, race, and key 
variables such as median income, housing affordability, rehospitalization rates, food stamps usage, 
uninsured rates, and racial demographics. Understanding these correlations is vital in comprehending 
the socio-economic and racial disparities prevalent in the studied region.

The disparity in median income over the studied years, particularly the peak observed in 2020, 
signifies economic fluctuations. Notably, county 1 stands out with significantly higher income 
compared to other counties, indicating potential disparities in resource access and opportunities. 
This wealth gap implies potential repercussions on community health outcomes. The consistent trend in 
the percentage of income spent on rent underscores a persistent challenge for residents, especially 
noticeable in counties 1 and 5, indicating enduring economic pressure.

The results pertaining to housing defined by renters whose rent is 50 percent of their income, 
reveal noteworthy patterns with respect to social needs. Examining the data across the four-year span 
indicates a consistent trend, as no significant differences were observed. However, significant differences are 
revealed when comparing between counties. Counties 1 and 5 exhibited a statistically significant increase 
in the percentage of renters facing housing costs amounting to 50 percent of their income compared to 
counties 2, 7, and 3. Intriguingly, there was no statistically significant distinction between counties 
1 and 5, suggesting a commonality in the challenges faced by renters in these specific regions. These 
findings underscore the importance of localized interventions and policy considerations to address the 
distinct socio-economic dynamics influencing housing affordability across different counties.

Food access was assesses through investigating the proportion of the population involved in the use of Food Stamps programs.
AHRQ data reveals notable variations across different counties. Counties 8, 2, and 5 exhibit a statistically significant 
higher percentage of the population relying on Food Stamps, suggesting a potentially elevated level of economic 
vulnerability or socio-economic challenges in these areas. County 6 also demonstrates a substantial proportion 
of its population on Food Stamps, indicating a notable need for social assistance. Conversely, County 7 stands 
out with the lowest percentage of its population depending on Food Stamps, implying comparatively better 
economic conditions or potentially more effective social support systems. These findings underscore the 
importance of targeted interventions and resource allocation to address social needs, particularly in counties 
with higher reliance on Food Stamps. Further exploration into the underlying factors contributing to these variations 
is needed for informed policy development and community-specific interventions.

Rehospitalization rates demonstrate stability across counties and years leading up to 2020, suggesting 
a consistent healthcare landscape. Counties 8, 2, and 5 exhibiting higher food stamp usage point to economic 
challenges faced by residents in these regions. This trend aligns with housing affordability issues, 
indicating a correlation between financial stress and reliance on government assistance programs.

The notably high uninsured rate in county 1 raises concerns about healthcare accessibility, likely 
linked to economic factors impacting insurance affordability. Conversely, the lower uninsured rate in 
Year 3 (2019) signifies positive progress. Analyzing the policies or interventions implemented during 
this period could provide valuable insights for effective healthcare reforms, offering potential guidance 
for future initiatives.

The racial demographics highlight disparities in various counties. County 3's higher percentage of white 
individuals, coupled with County 2, 5, and 1 having the highest black population, points to the need for
targeted interventions addressing racial health inequalities. These disparities might be indicative of 
historical, social, and economic factors influencing the healthcare experiences of different racial groups.

\subsection{Limitations}
One limitation lies in the availability and quality of data. This dataset does not have any 
data from years after 2020 which may serve to limit potential external validity considerations. 
There may also be variability in data collection methods and discrepancies in reporting standards 
leading to missing or incomplete data over the course of 4 years. Another limitation involves the scope 
of the study, focusing on specific counties in Connecticut may not fully capture nationwide disparities. 
Additionally, the research is limited to the factor parameters selected to investigate which  might not 
encompass all relevant social determinants affecting health outcomes.

\subsection{Future Directions}
These findings underscore the complex interplay between socio-economic status, race, and health outcomes. 
Addressing these disparities requires multifaceted interventions, including economic support, affordable 
housing initiatives, and targeted healthcare access programs. Future research should delve deeper into the 
root causes of these disparities, considering historical and systemic factors. Additionally, policy interventions 
and community-based programs should be designed to specifically target areas and populations facing the most 
significant challenges, aiming for a more equitable healthcare landscape for all residents.

\appendix

\bibliography{../manuscript/ref}
\bibliographystyle{chicago}

\end{document}

{housing by income.pdf}
{housing by year.pdf}
{Income by town.pdf}
{Inome by year.pdf}
{Rehosp by income/year.pdf}
{Stamps by income/year.pdf}